---
title: "Tracking Counties Visualizations"
author: "Evan Canfield"
date: "5/11/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE, 
                      warning=FALSE)
```


The following libraries were used.

```{r library_calls, message=FALSE, warning=FALSE}
#Library Calls
if (!require(pacman)) {install.packages('pacman')} 
library(pacman)

p_load(broom, grid, gridExtra, ggthemes, leaflet, RColorBrewer, tidyverse, tigris)

options(tigris_use_cache = TRUE)
```

#Map of States Visited

##Inputs
```{r state_shapefile_import, cache=TRUE, echo=TRUE, results='hide'}
state.sp <- states(cb = TRUE)
```

```{r blank_state_list}
state_list_blank<- state.sp@data %>% 
  select(NAME, STUSPS, GEOID) %>% 
  distinct() %>% 
  mutate(Visited = 0) %>%
  #Remove US Territories
  filter(GEOID < 60) %>% 
  arrange((GEOID))
```

##Shapefile Processing

```{r state_shapefile_processing, message = FALSE}
#State
state.sp@data$id <- rownames(state.sp@data)

state_points <- tidy(state.sp)

state.df <- state_points %>%
  left_join(y = state.sp@data, by = "id") %>% 
  filter(!(STATEFP %in% c("02", "15", "60", "66", "69", "72", "78"))) %>% 
  filter(long < 1)
```


##List of States Visited

```{r state_list_creation}
#Blank State List Export
# state_list_blank %>% 
#   write.csv("./data/list_of_states_blank.csv")

#Active State List Import
state_list_active <- read.csv("./data/list_of_states_active.csv", 
                              stringsAsFactors = FALSE)
```

###Join

```{r state_df_join}
state.df_visited <-left_join(state.df, state_list_active, by = c("STUSPS"))
```

##Plotting the Map

```{r color_fill}
cols <- c("#F4F6F6", "#1F618D")
```

```{r us_state_plot}
US_state <- ggplot(data = state.df_visited, 
               mapping = aes(x = long, y = lat, 
                             fill = factor(Visited), group = group))+
  geom_polygon(color = "gray80", size = 0.05) +
  scale_fill_manual(values = cols) +
  theme_map() + 
  guides(fill=FALSE) +
  coord_map(projection = "albers", lat0 = 39, lat1 = 45) +
  labs(title = "United States")

US_state
```


#Map of Counties Visited

To create a map of counties visited I more or less repeated the same steps performed to generate the state visual, only this time using county level data.  Again, I turned to the Tigris package.

##Inputs
```{r county_shapefile_import,  cache=TRUE, echo=TRUE, results='hide'}
#US County Longitude/Latitiude Informaiton from socviz Package
county.sp <- counties(cb = TRUE)
```

In order to track which counties I have been to I also needed document listing each of the counties with a column indicating whether or not I had visited. I generated that list from Tigris as well, using it's database of FIPS codes. FIPS codes are Federal Information Processing Standard Publication 6-4, a set of 5 digits which uniquely identifies each county or county-like equivalent in the United States. As FIPS codes are unique to each county they will act as good tracking identifier.

```{r county_list_blank}
county_list_blank <- fips_codes
```

In order to create an active list of counties I've visited I exported the blank list of counties, with an additional column used to indicate whether I've visited or not.

I filled that list out manually and imported it back into this analysis.
```{r county_list_active}
#Export Blank List of Counites
# county_list_blank_export <- county_list_blank %>% 
#   mutate(Visited = 0) %>% 
#   write.csv("./data/list_of_counties_blank.csv")

county_list_active_import <- read.csv("./data/list_of_counties_active.csv",
                                      stringsAsFactors = FALSE)

```

## Shapefile Processing
Like with the state latitude and longitude data, ggplot does not accept SpatialPolygonsDataFrame files. Therefore, I converted the file to a standard dataframe.

```{r, county shapefile conversion, warning=FALSE}
county.sp@data$id <- rownames(county.sp@data)

county_points <- tidy(county.sp, region = "id")

county.df <- county_points %>% 
  left_join(y = county.sp@data, by = "id") %>% 
  filter(!(STATEFP %in% c("02", "15", "60", "66", "69", "72", "78"))) %>% 
  filter(long < 1)
```

##List of Counties Visited
Now that the spatial information is in dataframe form and I have a completed Counties Visited document, also in dataframe form, I joined these files to create the final dataframe which to be used with ggplot. 

```{r county_df_processing}
#Prepare List of Blank Counties
county_list_blank_join <- county_list_blank %>%  
  mutate(state_code_num = as.numeric(state_code)) %>% 
  mutate(county_code_num = as.numeric(county_code)) %>% 
  select(state_code, county_code, state_code_num, county_code_num)

#Prepare Active List of Counties
county_list_active <- county_list_active_import %>% 
  left_join(y = county_list_blank_join, 
            by = c("state_code" = "state_code_num", 
                   "county_code" = "county_code_num")) %>%
  select(-X, -state_code, -county_code) %>% 
  rename("state_code" = "state_code.y", "county_code" = "county_code.y") %>% 
  select(state_code, county_code, everything()) %>% 
  mutate(FIPS = paste(state_code, county_code, sep = ""))

#Join County Latitude/Longitude Data with FIPS County Information - Dataframe
county.df_visited <- left_join(county.df, county_list_active, 
                               by = c("GEOID" = "FIPS"))
```

##Plotting the Map
I used the state latitude and longitude dataframe in this plot, as well as the county data. Overlaying the state boundaries data on the county data makes for a clearer map.

```{r us_county_plot}
US_county <- ggplot(mapping = aes(x = long, y = lat, group = group)) +
  geom_polygon(data = county.df_visited, 
                            mapping = aes(fill = factor(Visited)), 
                            color = "gray80", size = 0.05) +
  scale_fill_manual(values = cols) + 
  geom_polygon(data = state.df,
               color = "black", 
               fill = NA, 
               size = 0.1) +
  theme_map() + 
  guides(fill=FALSE) +
  coord_map(projection = "albers", lat0 = 39, lat1 = 45) +
  labs(title = "United States")
  
US_county
```

Well, that certainly looks a lot different than the map of just states visited. Let's take a look at the state level and county level maps side by side.

```{r us_state_and_county_plots}
grid.arrange(US_state, US_county, ncol = 2)
```

I need to get out more.

I am also particularly curious of the New England region as well as North Carolina. I grew up in Connecticut and my family lived all over New England. Family trips covered a lot of ground in those states. I currently live in North Carolina, and have for over 8 years. How much of these areas, in which I have spent so much of my life, have I really seen.

```{r new_england_county_plot}
#New England
county.df_visited_new_england <- county.df_visited %>% 
  filter(state %in% c("CT", "RI", "MA", "VT", "NH", "ME"))

state.df_new_england <- state.df %>%
  filter(STUSPS %in% c("CT", "RI", "MA", "VT", "NH", "ME"))


New_England_county <- ggplot(data = county.df_visited_new_england, 
              aes(x = long, y = lat, fill = factor(Visited), group = group)) +
  geom_polygon(color = "gray80", size = 0.05) +
  scale_fill_manual(values = cols) +
  geom_polygon(data = state.df_new_england,
               color = "black", 
               fill = NA, 
               size = 0.25) + 
  theme_map() + 
  guides(fill=FALSE) +
  coord_equal()+
  labs(title = "New England")

New_England_county
```

```{r north_carolina_county_plot}
#North Carolina
county.df_visited_nc <- county.df_visited %>% 
  filter(state == "NC")

state.df_nc <- state.df %>%
  filter(STUSPS %in% c("NC"))

NC_county <- ggplot(data = county.df_visited_nc, 
               aes(x = long, y = lat, fill = factor(Visited), group = group)) + 
  geom_polygon(color = "gray80", size = 0.05) + 
  scale_fill_manual(values = cols) +
  geom_polygon(data = state.df_nc,
                      color = "black", 
                      fill = NA, 
                      size = 0.25) +
  theme_map() + 
  guides(fill=FALSE) +
  coord_equal() +
  labs(title = "North Carolina")

NC_county
```

Honestly, no surprises here. For New England it's all of Connecticut, Rhode Island, and New Hampshire, with only Martha's Vineyard and Nantucket missed in Massachusetts, and one county being missed in northeast Vermont. But I have barely scratched the surface of Maine. Based on where my extended family resided when I was growing up, this map pretty much sums up where we would have traveled for family visits and vacations.

For North Carolina, I have been everywhere in the western part of the state. I currently live in Charlotte, which is close-ish to the center of the state. Since I've always been a bigger fan of the mountains than the beach, heading west from Charlotte more than east tracks. 

#Leaflet

To try one last thing, I want to generate an interactive visual. I used Leaflet for this. I plant on mapping the same shaded polygons I previously plotted in ggplot on top of the leaflet map. For this plot we do not need to filter out US Territories or Alaska or Hawaii. A complete map of the world is generated regardless. 

An additional wrinkle to using Leaflet is that unlike ggplot, leaflet can use SpatialPolygonsDataFrames. Therefore, we need to create a new dataset that incorporates the county SpatialPolygonsDataFrame along with the Counties Visited information.

```{r county_visited_geospatial}
#Join County Latitude/Longitude Data with FIPS County Information - Geo Spatial
county.sp_visited <- geo_join(spatial_data = county.sp, 
                              data_frame = county_list_active,
                              by_sp = "GEOID", by_df = "FIPS")
```

##Leaflet Visual
Now that a final dataset is available I plotted it in leaflet. Just like in the ggplot county visual I also plot the state SpatialPolygonsDataFrame in order to better visualize the state lines. 

```{r leaflet}

pal <- colorNumeric(palette = "viridis", domain = county.sp_visited$Visited)

leaflet(options = leafletOptions(minZoom = 3)) %>% 
  addTiles() %>% 
  addPolygons(data = state.sp,
    color = "#2E4053",
    weight = 1,
    opacity = 1,
    dashArray = 1.5,
    fillColor = "white",
    fillOpacity = 0
  ) %>%
  addPolygons(data = county.sp_visited,
    highlight = highlightOptions(color = "white"),
    fillColor = ~pal(county.sp_visited$Visited),
    weight = 0.5,
    opacity = 1,
    color = "white",
    dashArray = 2.5,
    fillOpacity = 0.25,
    label =  ~paste(NAME)
  ) %>% 
  setView(lng = -96, 
          lat = 37.8, 
          zoom = 4) %>%
  setMaxBounds(lng1 = -180, 
               lng2 = -60,
               lat1 = 73, 
               lat2 = 15)
```

This leaflet visual runs slower that I would ideally want. In the future I hope to figure out a way to get the app to run more smoothly. But this interactive visual is not just another way to visualize  counties I been to. It is perhaps more useful as a tool to determine which counties I may try to visit on future trips. Roads, towns, and other places of interest are very easy to see in relation to county lines.